---
title: ''
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rgdal)
library(tidyverse)

#set downloads path for whoever is downloading the source file (see notes below)
downloads_path <- keyring::key_get('downloads_path')

```

Source for the municipality shapefile is the Cook County GIS Department. The file is available online through the GIS Department's Open Data Portal [here](https://hub-cookcountyil.opendata.arcgis.com/datasets/534226c6b1034985aca1e14a2eb234af_2/explore?location=41.811250%2C-87.890600%2C10.56). The shapefile has been exported from this site and imported below for cleaning.

*Note: In future versions of this script, manual exporting of the shapefile could potentially be eliminated. For future iterations, explore importing the file via geojson or the arcgisbindings package.*

```{r import}

#import downloaded source file from GIS dept
raw_munis <- readOGR(dsn = paste0(downloads_path, "Municipality"), layer = "Municipality")

#import CCDPH standard municipality data set
muni_data <- read_csv("https://github.com/Cook-County-Department-of-Public-Health/ccdph-data-sets/blob/main/Municipality%20Populations.csv?raw=TRUE")

```

Create custom fields for CCDPH use

```{r clean}

#make sure all municipality names match between shapefile and data set and all towns are accounted for in both datasets
raw_munis@data[!(raw_munis@data$MUNICIPALI %in% muni_data$Municipality) & !is.na("raw_munis@data$MUNICIPALI"), "MUNICIPALI"]
muni_data[!(muni_data$Municipality %in% raw_munis@data$MUNICIPALI), ]

#Fix McCook Spelling
mccook_index = which(raw_munis@data$MUNICIPALI == "Mc Cook")
raw_munis@data$MUNICIPALI = as.character(raw_munis@data$MUNICIPALI)
raw_munis@data$MUNICIPALI[mccook_index] = "McCook"

#create fresh copy to modify
munis_customized <- raw_munis 

#link to muni data and create variables  
munis_customized@data <- munis_customized@data %>%
  #rename_with(~paste0(.x, "_gis_dept")) %>%
  left_join(muni_data, by = c("MUNICIPALI" = "Municipality")) %>%
  janitor::clean_names() %>%
  select(muni_name = municipali,
         pop_2010 = population2010,
         district,
         partial,
         exclude = exclude_from_analysis,
         fips = census_place_code,
         gis_agncy = agency, 
         gis_name = agency_des,
         gis_area = shapest_are,
         gis_leng = shapest_len
         )

```

Export customized shapefile in state plane and WGS projections. State plane projection was verified in ArcPro by overlaying with GIS Dept raw muni file.

```{r export}

### STATE PLANE FILE ###

#Create directory to hold shapefile
stateplane_path <- paste0(getwd(), "/epsg-3435-illinois-stateplane-east/ccdph-cook-county-munis")
dir.create(path = stateplane_path)

#write shapefile
writeOGR(obj=munis_customized, dsn = stateplane_path, layer = "ccdph-cook-county-munis", driver = "ESRI Shapefile")

#zip shapefile
zip(zipfile = stateplane_path, files = dir(stateplane_path, full.names = TRUE))

#remove unzipped files
unlink(stateplane_path, recursive = T)


### WGS FILE ###

#Transform file from state plane to WGS
munis_customized_WGS <- spTransform(munis_customized, CRS("+init=epsg:4326"))

#Create directory to hold shapefile
wgs_path <- paste0(getwd(), "/epsg-4326-wgs84-for-leaflet/ccdph-cook-county-munis-wgs")
dir.create(path = wgs_path)

#write shapefile
writeOGR(obj=munis_customized_WGS, dsn = wgs_path, layer = "ccdph-cook-county-munis-wgs", driver = "ESRI Shapefile")

#zip shapefile
zip(zipfile = wgs_path, files = dir(wgs_path, full.names = TRUE))

#remove unzipped files
unlink(wgs_path, recursive = T)

```


